Extended Data

### 2022-03-21 doc/manuscript/Dow_manuscript_Nature.Rmd
acknowledge reviewers
Update Dow_manuscript_Nature.Rmd

Dendrometer band measurements were collected at SCBI[@bourg_initial_2013] and Harvard Forest[@finzi_carbon_2020;@dorangeville_peak_2022], both part of the Forest Global Earth Observatory (ForestGEO)[@anderson-teixeira_ctfsforestgeo_2015;@davies_forestgeo_2021]. 
SCBI (38.8935° N, 78.1454° W; elevation 273–338 m.a.s.l.) is located in the Blue Ridge Mountains at the northern end of Shenandoah National Park, 5 km south of Front Royal, Virginia. 
The forest is secondary and mixed age, having established in the mid-19th century after conversion from agricultural fields[@bourg_initial_2013]. 
Dominant canopy species within the 25.6 ha ForestGEO plot include tulip poplar (*Liriodendron tulipifera* L.), oaks (*Quercus spp.*), and hickories (*Carya spp.*)[@helcoski_growing_2019]. 
The climate is humid temperate, with 1950-2019 mean annual precipitation of 1018 mm and temperatures averaging 1°C in January and 24°C in July[@anderson-teixeira_joint_2022].
Within the study period for the dendrometer band analysis (2011-2020), Spring temperatures (March + April $T_{max}$) and summer SPEI values (4-month value of August) were similar to the average prior to the study period (1970-2010). Prior to the study period, the average spring $T_{max}$ was 16.9 $degreeC$ and summer SPEI was -0.06. In the study period, average spring $T_{max}$ was 17.6 $degreeC$ and summer SPEI was 0.51.

Harvard Forest (42.5388° N, 72.1755° W, 340-368 m.a.s.l.) is located near the central Massachusetts town of Petersham. 
The forest is secondary and mixed age, having re-established around the beginning of the 20th century following agricultural use and significant hurricane damage in 1938. 
Dominant species within the 35 ha ForestGEO plot are hemlock (*Tsuga canadensis* (L.) Carrière), oak (*Quercus spp.*) and red maple (*Acer rubrum* L.). 
The climate is temperate continental, with 1950-2019 mean annual precipitation of 1104 mm and temperatures averaging -5° C in January and 22° C in July[@anderson-teixeira_joint_2022]. 
Within the study period for the dendrometer band analysis (1999-2003), Spring temperatures (March + April $T_{max}$) and summer SPEI values (4-month value of August) were similar to the average prior to the study period (1970-1998). Prior to the study period, the average spring $T_{max}$ was 10.9 $degreeC$ and summer SPEI was 0.13. In the study period, average spring $T_{max}$ was 11.2 $degreeC$ and summer SPEI was 0.20.
The summer of 1999 was abnormally dry, with precipitation of only 52 mm mo^-1^ in June-August compared to average monthly precipitation of 100+ mm[@dorangeville_peak_2022].

@@

# Acknowledgements {-#Acknowledgements} 
We gratefully acknowledge all researchers who assisted with data collection in the field and laboratory, particularly Tsun Fung Au, Joshua Bregy, James Dickens, Karen Heeter, Anna Hennage, Daniel King, James McGee, Benjamin Lockwood, Jennifer McGarvey, Victoria Meakem, Josh Oliver, Jessica Shue, Karly Schmidt-Simard, Brandon Strange, Alyssa Terrell, Brynn Taylor, Michael Thornton, Senna Robeson, Matt Wenzel, and Luke Wylie. 
Thanks to David A. Orwig and members of the ForestGEO Ecosystems & Climate Lab at SCBI for helpful feedback.
The research was funded by ForestGEO (Smithsonian). 
Collection of tree-ring samples was funded by a USDA Agriculture and Food Research Initiative grant 2017-67013-26191 and from the Indiana University Vice Provost for Research Faculty Research Program.

>>>>

Dendrometer band measurements were collected at SCBI[@bourg_initial_2013] and Harvard Forest[@finzi_carbon_2020;@dorangeville_peak_2022], both part of the Forest Global Earth Observatory (ForestGEO)[@anderson-teixeira_ctfsforestgeo_2015;@davies_forestgeo_2021]. 
SCBI (38.8935° N, 78.1454° W; elevation 273–338 m.a.s.l.) is located in the Blue Ridge Mountains at the northern end of Shenandoah National Park, 5 km south of Front Royal, Virginia. 
The forest is secondary and mixed age, having established in the mid-19th century after conversion from agricultural fields[@bourg_initial_2013]. 
Dominant canopy species within the 25.6 ha ForestGEO plot include tulip poplar (*Liriodendron tulipifera* L.), oaks (*Quercus spp.*), and hickories (*Carya spp.*)[@helcoski_growing_2019]. 
The climate is humid temperate, with 1950-2019 mean annual precipitation of 1018 mm and temperatures averaging 1°C in January and 24°C in July[@anderson-teixeira_joint_2022].
Within the study period for the dendrometer band analysis (2011-2020), spring temperatures (March + April $T_{max}$) and summer SPEI values (4-month value of August) were similar to the average prior to the study period (1970-2010). Prior to the study period, the average spring $T_{max}$ was 16.9 $degreeC$ and summer SPEI was -0.06. In the study period, average spring $T_{max}$ was 17.6 $degreeC$ and summer SPEI was 0.51.

Harvard Forest (42.5388° N, 72.1755° W, 340-368 m.a.s.l.) is located near the central Massachusetts town of Petersham. 
The forest is secondary and mixed age, having re-established around the beginning of the 20th century following agricultural use and significant hurricane damage in 1938. 
Dominant species within the 35 ha ForestGEO plot are hemlock (*Tsuga canadensis* (L.) Carrière), oak (*Quercus spp.*) and red maple (*Acer rubrum* L.). 
The climate is temperate continental, with 1950-2019 mean annual precipitation of 1104 mm and temperatures averaging -5° C in January and 22° C in July[@anderson-teixeira_joint_2022]. 
Within the study period for the dendrometer band analysis (1999-2003), spring temperatures (March + April $T_{max}$) and summer SPEI values (4-month value of August) were similar to the average prior to the study period (1970-1998). Prior to the study period, the average spring $T_{max}$ was 10.9 $degreeC$ and summer SPEI was 0.13. In the study period, average spring $T_{max}$ was 11.2 $degreeC$ and summer SPEI was 0.20.
The summer of 1999 was abnormally dry, with precipitation of only 52 mm mo^-1^ in June-August compared to average monthly precipitation of 100+ mm[@dorangeville_peak_2022].

@@

# Acknowledgements {-#Acknowledgements} 
We gratefully acknowledge all researchers who assisted with data collection in the field and laboratory, particularly Tsun Fung Au, Joshua Bregy, James Dickens, Karen Heeter, Anna Hennage, Daniel King, James McGee, Benjamin Lockwood, Jennifer McGarvey, Victoria Meakem, Josh Oliver, Jessica Shue, Karly Schmidt-Simard, Brandon Strange, Alyssa Terrell, Brynn Taylor, Michael Thornton, Senna Robeson, Matt Wenzel, and Luke Wylie. 
Thanks to David A. Orwig, members of the ForestGEO Ecosystems & Climate Lab at SCBI, and four anonymous reviewers for helpful feedback.
The research was funded by ForestGEO (Smithsonian). 
Collection of tree-ring samples was funded by a USDA Agriculture and Food Research Initiative grant 2017-67013-26191 and from the Indiana University Vice Provost for Research Faculty Research Program.
